Supporting Information for ``The diversity of cloud responses to twentieth-century sea surface temperatures"

### 2019-10-14 Silvers_etal_grl_SMrevisions_2017_sub.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified July 2014
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format: two column (default)
% and draft.
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%
% Choose the journal abbreviation for the journal you are
% submitting to:

% jgrga JOURNAL OF GEOPHYSICAL RESEARCH
% gbc   GLOBAL BIOCHEMICAL CYCLES
% grl   GEOPHYSICAL RESEARCH LETTERS
% pal   PALEOCEANOGRAPHY
% ras   RADIO SCIENCE
% rog   REVIEWS OF GEOPHYSICS
% tec   TECTONICS
% wrr   WATER RESOURCES RESEARCH
% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw    SPACE WEATHER
% ms    JAMES
% ef    EARTH'S FUTURE
% ea    EARTH AND SPACE SCIENCE
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

%\documentclass[draft,grl]{agutex2015}
\documentclass[grl]{agutex2015}
% To create numbered lines:

% If you don't already have lineno.sty, you can download it from
% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
% (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
% Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno}
% and \linenumbers*[1]command, below:

\usepackage{graphicx}
\usepackage{epstopdf}

% \usepackage{lineno}
% \linenumbers*[1]
%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
%
% DO NOT USE \psfrag or \subfigure commands.
%
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
%  \usepackage[dvips]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
  \setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

@@

% Shorter version of title entered in capital letters:
\titlerunninghead{Cloud Feedbacks in long-AMIP}

%Corresponding author mailing address and e-mail address:
%\authoraddr{Corresponding author: A. B. Smith,
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(a.b.smith@hwr.arizona.edu)}

@@

\title{Decadal variations of cloud feedback in long-AMIP experiments with three GFDL climate models.}
%
% e.g., \title{Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
%

@@

% \altaffilmark will produce footnote;
% matching \altaffiltext will appear at bottom of page.
\authors{Levi G. Silvers, David Paynter, and Ming Zhao}
% \authors{A. B. Smith,\altaffilmark{1}
% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}

@@

%% ------------------------------------------------------------------------ %%
%
%  KEYPOINTS
%
%% ------------------------------------------------------------------------ %%

% Key points are 1 to 3 points that the author provides,
% that are 100 characters or less, that are ultimately published
% with the article.
%% for example:
% \keypoints{\item Here is the first keypoint. what happens if it is a
% long keypoint, like this one. We want to see this wrap please.
% \item This is the second.
% \item And here is the third keypoint
% }

\keypoints{\item Three models show the Eastern Tropical Pacific as a key location where decadal trends in SST and low-level clouds strongly influence the global energy budget
\item Changes in low-level clouds can be linearly decomposed but with differing magnitudes of dependencies on EIS and SST
\item The cloudy response of the models differs markedly between twentieth century periods with high and low climate sensitivities}
%\item AM2.1, AM3, and AM4g10r8 all show strong twentieth
%century decadal variability of the global feedback parameter.
%\item All three models show increased low-level clouds in the East Pacific
%over recent decades.
%\item Low-level cloud trends depend on both EIS trends 
%and omega trends.}

%% Keypoints will print underneath the abstract.

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
%% ------------------------------------------------------------------------ %%

% >> Do NOT include any \begin...\end commands within
% >> the body of the abstract.

\begin{abstract}

Significant decadal variability of the global feedback parameter is shown to be clearly linked to the pattern of surface warming through the evolution of low-clouds and the estimated inversion strength (EIS).  We examine the atmospheric response to changing SST patterns between 1870-2004 using the amip-piForcing experimental configuration with three GFDL atmospheric global climate models (AM2, AM3, and AM4).  There is a notable strengthening of the negative global feedback from the middle period of the 20th century until the present.  We contrast the variability of the cloud response among the models between these two periods.  The changing low-cloud cover is analyzed within a framework that includes trends of the EIS and the mid-tropospheric vertical velocity.

Recently it has been argued that decadal changes in low-level cloud cover over the eastern Pacific ocean basin strongly influence the global net feedback parameter.  These decadal changes of low-level cloud cover have been tied to corresponding changes in the observed sea surface temperatures.  We examine these claims with AM2, AM3, and AM4.  The dominant configuration difference among the models is the parameterization of the convective clouds.  The amip-piForcing experiment provides an ideal configuration to test the range of atmospheric responses to a prescribed but evolving SST pattern.

All three GFDL models confirm the importance of the relationship between the global climate sensitivity and the trends in the eastern Pacific of SST and low-level clouds.  However, there are noteworthy differences in the cloud response of the models.  Despite all three models being driven by observed SSTs, sea ice concentration, and constant preindustrial forcing agents distinct net global feedbacks are simulated.

%Recently it has been argued that decadal changes in low-level cloud cover over the eastern Pacific ocean basin strongly influence the global net feedback parameter.  These decadal changes of low-level cloud cover have been tied to corresponding changes in the observed sea surface temperatures (SST).  We examine this claim in three generations of GFDL atmospheric global climate models $(AM2.1, AM3, and AM4_g10r8)$.  The dominant configuration difference among the models is the parameterization of convective clouds.  Atmospheric experiments with observed SSTs (atmospheric forcing held at constant preindustrial values) between 1870-2004 are examined to elucidate the source of differences between the models.  
%
%All three models are shown to have strong decadal variability in their net feedback parameter, the global mean cloud-radiative effect, and the low-level cloud fields.  All three models confirm the importance of the relationship in the eastern Pacific region between decadal SST patterns and low-level cloud changes.  However, there are noteworthy differences in the cloud response of the models.  
%
%The correlation between changes in low-level cloud cover and the lower tropospheric stability is used to examine the impact of the alternate cumulus parameterizations used by the models.  We approximate the change of low-level cloud with a linear combination of changes in lower tropospheric stability and SST.  The usefulness of this approximation differs among the three models examined.   
%
%Despite all models being driven by observed SSTs, sea ice concentration, and having constant preindustrial atmospheric forcing agents distinct net global feedbacks are simulated. 
%
\end{abstract}

@@

\section{Introduction}

The global energy budget of Earth is often considered in a mean, equilibrated context.  However, fluctuations in the TOA net radiative flux are observed on monthly, yearly, and decadal scales.  The decadal variability of the energy budget is also present in the global mean cloud radiative effect.  This paper elucidates the nature of changes to the cloud fields in simulations covering 1871-2005.  By clarifying key geographical regions and particular types of clouds that play a prominent role in the decadal variability of the TOA net radiative flux we hope to increase our confidence in the equilibrated energy budget analysis of Earth.

    \begin{enumerate}
        \item Decadal variations of global feedback in AMIP-like experiments offer an opportunity for 
        mechanistic attribution
        	\begin{enumerate}
	    \item SST patterns have been shown to be important for variations of global 
	    feedback (Gregory and Andrews, 2016).  
	    \item Estimated Inversion Strength (EIS)

	    The lower-tropospheric stability, and the related estimated inversion strength (EIS) has been used as a predictor of low-level clouds in select stratocumulus dominated regions (Klein and Hartmann, 1993, Wood and Hartmann, 2006).  The tropical marine EIS was found by Zhou et al. 2016 to be well correlated with the evolution of tropical tropical low-cloud change.  The EIS can be approximated by a linear combination of the surface temperature and the temperature on the 700hPa pressure level.  In the tropics, this provides a glimpse of one influence of the convection parameterization because the mid-tropospheric temperature is largely determined by the lapse rate in the warm pool regions with strong deep convection.

	    \item Analyzing alternate atmospheric climate models with the same SST 
	    patterns and atmospheric forcing throughout the twentieth century highlights
	    both atmospheric internal variability and the impact of different parameterization 
	    configurations among the models.
	    \item Why AM2.1, AM3, and AM4g10r8?
	\end{enumerate}
	\item Relationship between decadal feedback, equilibrium climate sensitivity, and effective climate sensitivity
	\item Review of Gregory and Andrews 2016, Zhou et al. 2016, Qu et al. 2014,2015, and Paytner et al. 2017
    \end{enumerate}

\section{Method}
%    \begin{enumerate}
 %       \item Description of amipPiForcing experiments
\subsection{Description of amipPiForcing experiments}

        The primary experiment used in this work corresponds to the amip-piForcing experiment.  For this experiment the AGCMs are run using observed SST and sea-ice, but with the forcing due to GHG, aerosol, and solar forcing set to assumed pre-industrial levels and held constant throughout the experiment.  The time simulated is from 1870 to 2005.  The amip-piForcing experiments are sometimes referred to as a `long-AMIP' experiments.  The amip-piForcing experiment has been included as one of the tier-II experiments for CFMIP3 (Webb et al. 2017), for additional details and examples see Gregory and Andrews (2016) and Zhou et al. (2016).

\subsection{Description of AM2.1, AM3, and AM4g10r8}

        The three GFDL atmospheric models analyzed in this paper have many differences, but the those that are thought to be of primary importance to the top-of-the-atmosphere radiation budget are relatively few.  The AM3 and AM4g10r8 models include an aerosol indirect effect by including the cloud drop number as a prognostic variable.  The AM2.1 model diagnosed the cloud drop number.  The parameterization schemes used to represent cumulus convection differs significantly among the three models.  The AM2.1 model uses the Relaxed-Arakawa-Schubert (RAS) scheme, AM3 uses the Donne-Deep scheme for deep convection and the University of Washington shallow convection scheme (UWShCu, Bretherton et al. 2004) for shallow convection.  AM4g10r8 uses a version of the UWShCu scheme that has been modified to include two bulk plumes which can account for both shallow and deep convection.  In addition to these differences, the scheme which computes the radiative transfer was updated from SEALW99/ESFSW99 in AM2.1 and AM3 to SEALW16/ESFSW16 in AM4g10r8.

\subsection{Methodology of Analysis (here or in results?)}        
%        \begin{enumerate}
%            \item Global feedback parameter via Gregory regression sliding window
            Because our interest is in the covariation of R and T rather than the change relative to a control state we compute the differential climate feedback parameter in the same was as in Gregory and Andrews [2016].  That is, linear regression is used to compute the trend between R and T.  The regression is applied to the global yearly mean values over 30 windows.  The values plotted in Figure 1 represent the mid-point of each window.  
 %           \item Global maps and global mean time series of trends per years
            To determine the spatial pattern of changes to particular variables linear regression was used at each grid point.              
 %           \item Estimated Inversion Strength

            The estimated inversion strength can be thought of as a correction to the LTS based on the moist adiabatic profile.  The LTS is defined to be $\rm{LTS}=\theta_{700}-\theta_{sfc}$.  
            To compute the estimated inversion strength (EIS) we follow the approximation made by Wood and Bretherton [2006]: 
            \begin{equation}
              \rm{EIS} = \rm{LTS} - \Gamma^{850}_m (z_{700}-\rm{LCL})
            \end{equation}
            with the lifting condensation level (LCL) calculated as $\rm{LCL}=(20+(T_{sfc}-273.15)/5)(1-RH_{sfc})$.
            The potential temperature gradient for moist-adiabatic processes is calculated as
            \begin{equation}
              \Gamma_m(T,p)=\frac{g}{c_p} \left[ 1-\frac{1+L_v q_s(T,p)/R_a T}{1+L^2_v q_s(T,p)/c_p R_v T^2} \right].
            \end{equation}
            To compute $\Gamma^{850}_m$ we used the temperature on the $850 \rm{hPa}$ pressure level and computed the saturation vapor pressure on the 
            $850 \rm{hPa}$ level using $q_s=610.8 \, \rm{exp}[17.27(T_{850}-273.15)/(T_{850}-35.85)] (\rm{hPa})$.  When the pressure surfaces of interest for these calculations intersect topography 
            the variables we are interested in are not defined.  To account for this, missing vallues of $\rm{T}_{850}$ are set to 50 $K$, $\rm{T}_{sfc}$ is not allowed to be lower than $200$ K, and $\rm{RH}_{sfc}$
            is not allowed to be lower than $5 \%$.

%    \end{enumerate}

\section{Results}

%\begin{enumerate}
 %   \item General Results
\subsection{General Results}

We have shown that all three models exhibit strong decadal variability of the feedback parameter.  The close connection between the global feedback and the CREs is clear based on their respective time series.  Correlations between the global feedback parameter and CRE for each model are: 0.93, 0.93, and 0.81 for AM2.1, AM3, and AM4g10r8, respectively.

Over the second half of the twentieth century AM4g10r8 and AM3 show a strong decrease of the global feedback parameter and the CRE response (Fig. \ref{fig:alphacre}).  Both of these models also have an increasing trend of EIS during this time period (Fig. \ref{fig:alphacre}).  In contrast, AM2.1 has a weak decrease of global feedback parameter and a nearly constant CRE trend during this time period.  The trend of EIS for AM2.1 is relatively flat and less than that for AM4g10r8 and AM3.  
%If the change of LLC can be linearly decomposed into a component due to the EIS and one due to the change of SST I would then expect this to indicate a poorer correlation for AM2.1 compared to AM3 in the tropics.  
Between 1925 and 1955 the average global feedback parameter for AM4, AM3, and AM2.1 changes from average values of (-1.35, -1.02, -1.80) to (-2.22, -2.42, -2.20) between 1975 and 1989.  This represents a dramatic change in the global feedback parameter and determining the cause will be an important part of our ability to understand how clouds influence the global energy budget on decadal time scales.

%Changes in low-cloud cover where approximated by Zhou et al. 2016 as $\Delta \rm{LCC} = \delta \Delta \rm{EIS} - \gamma \Delta \rm{SST}$.  If this holds then to the extent that different models use the same SST fields, the change of LCC should follow the change of EIS.  
%The importance of different SST patterns with the same mean value has been recognized (Andrews et al. 2015).  Is there a similar importance of the EIS pattern?  In this context the hypothesis (Qu et al. 2014,2015) that changes of EIS can be approximated as 
%$\Delta \rm{EIS} \approx \alpha \Delta T_{700}-\beta \Delta T_s $ indicates that the pattern of EIS change is entirely (for the case of the same SST) determined by the change of temperature on the 700 hPa level.  This is an easier quantity to diagnose from GCMs than EIS.

%All three models show an increase of the anomalous tropical mean low-level clouds and EIS between about 1975 and 2005, with the increase in AM2.1 covering the smallest period of time.   
%The tropical marine low-cloud changes in AM4g10r8 are well explained (r=0.96) by a linear combination of the changes in the EIS and the SST.  
%To the extent that the MAC LWP from Norris et al. 2016 represents a realistic trend of LWP the AM4g10r8 LWP trend shows a more realistic trend than AM2.1.  
%Because of the large amount of the anomalous change in low-cloud cover that can be explained by the linear combination of EIS and SST, as in CAM5.3, it is tempting to conclude that AM4g10r8 is doing the best job at reproducing observed relationships between EIS and low-cloud cover.  
%However, AM4 needs to be tested for its representation of the cloud controlling factors which have been discussed by Bretherton, and are shown in Figure 1 of the supplementary material of Zhou et al. 2016.

The period after about 1975 is found by all three models to have a small climate sensitivity (large, negative feedback parameter, Fig. \ref{fig:alphacre}) and a negative CRE.  This period also corresponds in all three models to an increase of tropical mean low-level clouds  (Fig. \ref{fig:lccts}).  Computing the trend of LCC change between 1975-2005 reveals much of this change to be concentrated in the eastern Pacific ocean basins (Fig. \ref{fig:lcceis_late}) in agreement with the findings of Zhou et al. 2016.  Further analysis of the changes to low-level clouds is discussed in the next section.

A strong relationship between LCC and EIS in many regions of the globe is shown by the trends of LCC and EIS between 1975-2005 (Figure \ref{fig:lcceis_late}) and between 1925-1955 (Figure \ref{fig:lcceis_early}).  Both time periods have a large increase of both EIS and LCC in the equatorial eastern Pacific ocean.  However, during the later period there is an increase of LCC throughout much of the Southern Hemispheric oceans while the early period the LCC decreases in much of the Southern Hemispheric oceans and northern midlatitudes.  This is consistent with a positive cloud feedback/large climate sensitivity during the early period and a small cloud feedback/climate sensitivity during the later period (Fig. \ref{fig:alphacre}).

%    \item Multiple Linear Regression 
\subsection{Multiple Linear Regression}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified November 2013
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format: two column (default)
% and draft.
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%
% Choose the journal abbreviation for the journal you are
% submitting to:

% jgrga JOURNAL OF GEOPHYSICAL RESEARCH
% gbc   GLOBAL BIOCHEMICAL CYCLES
% grl   GEOPHYSICAL RESEARCH LETTERS
% pal   PALEOCEANOGRAPHY
% ras   RADIO SCIENCE
% rog   REVIEWS OF GEOPHYSICS
% tec   TECTONICS
% wrr   WATER RESOURCES RESEARCH
% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw    SPACE WEATHER
% ms    JAMES
% ef    EARTH'S FUTURE
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

\documentclass[draft,grl]{agutexSI}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  SUPPORTING INFORMATION TEMPLATE
%
%% ------------------------------------------------------------------------ %%
%
%
%Please use this template when formatting and submitting your Supporting Information.

%This template serves as both a “table of contents” for the supporting information for your article and as a summary of files.
%
%
%OVERVIEW
%
%Please note that all supporting information will be peer reviewed with your manuscript.
%In general, the purpose of the supporting information is to enable authors to provide and archive auxiliary information such as data %tables, method information, figures, video, or computer software, in digital formats so that other scientists can use it.
%The key criteria are that the data:
% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
%    including %data so the experiment can be reproducible);
% 2. are likely to be usable or used by other scientists working in the field;
% 3. are described with sufficient precision that other scientists can understand them, and
% 4. are not exe files.
%
%USING THIS TEMPLATE
%
%***All references should be included in the reference list of the main paper so that they can be indexed, linked, and counted as citations.  The reference section does not count toward length limits.
%
%All Supporting text and figures should be included in this document. Insert supporting information content into each appropriate section of the template. %Figures and tables should appear above each caption.  To add additional captions, simply copy and paste each sample caption as needed.

%Tables may be included, but can also be uploaded separately, especially if they are larger than 1 page, or if necessary for retaining table formatting. Data sets, large tables, movie files, and audio files should be uploaded separately, following AGU naming conventions. Include their captions in this document and list the file name with the caption. You will be prompted to upload these files on the Upload Files tab during the submission process, using file type “Supporting Information (SI)”

%IMPORTANT NOTE ON FIGURES AND TABLES
% Placeholders for figures and tables appear after the \end{article} command, after references.
% DO NOT USE \psfrag or \subfigure commands.
%
%  Uncomment the following command to include .eps files
% \usepackage[dvips]{graphicx}
 % \usepackage[dvipdf]{graphicx}
  \usepackage{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
\setkeys{Gin}{draft=false}
%\setkeys{Gin}{draft=true}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

@@

% Shorter version of title entered in capital letters:
%\titlerunninghead{SHORT TITLE}
\titlerunninghead{Cloud Feedbacks in long-AMIP}

%Corresponding author mailing address and e-mail address:
\authoraddr{Corresponding author: L. G. Silvers,
Princeton University/GFDL, Princeton, New Jersey, USA.
(silvers@princeton.edu)}
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(a.b.smith@hwr.arizona.edu)}

@@

%\includegraphics{agu_pubart-white_reduced.eps}

\title{Supporting Information for ``The diversity of cloud responses to twentieth-century sea surface temperatures"}
%
% e.g., \title{Supporting Information for "Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$"}
%
%DOI: 10.1002/%insert paper number here%

@@

% \altaffilmark will produce footnote;
% matching \altaffiltext will appear at bottom of page.

\authors{Levi G. Silvers\altaffilmark{1}, David Paynter\altaffilmark{2}, and Ming Zhao\altaffilmark{2}}

\altaffiltext{1}{Princeton University/GFDL, Princeton, New Jersey, USA}
\altaffiltext{2}{GFDL/NOAA, Princeton, New Jersey, USA}

% \authors{A. B. Smith,\altaffilmark{1}
% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}

@@

@@

\noindent\textbf{Contents of this file}
%%%Remove or add items as needed%%%
\begin{enumerate}
\item Additional Information for Boundary Condition and Parameterization details
\item Additional Results
\item Computational details of the Estimated Inversion Strength
\item Figures S1 to S4
\item Tables S1 to S2
%if Tables are larger than 1 page, upload as separate excel file
\end{enumerate}
%\noindent\textbf{Additional Supporting Information (Files uploaded separately)}
%\begin{enumerate}
%\item Captions for Datasets S1 to Sx
%\item Captions for large Tables S1 to Sx (if larger than 1 page, upload as separate excel file)
%\item Captions for Movies S1 to Sx
%\item Captions for Audio S1 to Sx
%\end{enumerate}

\noindent\textbf{Additional Information for Boundary Condition and Parameterization Details}
%Type or paste your text here. The introduction gives a brief overview of the supporting information. You should include information %about as many of the following as possible (when appropriate):
% 1. a general overview of the kind of data files;
% 2. information about when and how the data were collected or created;
% 3. a general description of processing steps used;
% 4. any known imperfections or anomalies in the data.
This experiment drives AGCMs with observationally based SST and sea-ice concentrations while the forcing due to greenhouse gases, aerosol, and solar forcing, is held constant at assumed pre-industrial levels.  However, due to the different time at which these models were actively used there are small differences in the boundary conditions. 
For the SST and sea ice concentration boundary conditions the AM2.1 model uses a combination of HADISST and NOAA data [Hurrel et al., 2008], AM3 uses version 1 of HADISST, and AM4.0 uses PCMDI-AMIP SST and sea ice boundary conditions which have been derived from observations, the method of derivation is described in Taylor et al. 2000 and Hurrel et al., 2008 with current details available at https://pcmdi.llnl.gov/mips/amip/.  The preindustrial forcing is based on assumed values from 1860 for AM2.1 and AM3, but from 1850 for AM4.0.  It is not expected that these differences in the initial conditions and boundary conditions would lead to alternate climate responses.  The land surface evolves freely.

The AM4.0 and AM3 models include an aerosol indirect effect with a prognostic cloud drop number.  The AM2.1 model specified the cloud drop number to be 100 (300) over ocean (land).

\noindent\textbf{Additional Results}

The paper by Qu et al., 2014 documented an inter-model analysis of the statistics from 5 geographic windows which were identified as important for low cloud cover.  Their windows encompass large swaths of the oceans which include many cloud types.  We have computed the regional decomposition of the climate feedback parameter over their 5 windows using the same methodology described in our primary manuscript (SPZ herafter; Silvers, Paynter, and Zhao).  Results are shown in Figure \ref{fig:quwindows}.  
This figure is identical to SPZ Figure 1.b, but with the climate feedback parameter computed with Qu et al.'s windows shown as the thin small-dashed lines.   
The rest of the tropics (between $\pm 30^{\circ}$ minus the 5 windows) are shown with thin long-dashed lines.  The conclusion that most of the decadal variability of the climate feedback parameter is due to the greater tropics and not due to the regions defined by these specific geographic windows holds true.  There is very little difference between the results of the windows used in the primary analysis and those from Qu et al., 2014.  This indicates that even if our region of interest is expanded, and the northern tropical Atlantic region included as in Qu et al., 2014, the temporal variability of $\alpha$ is still not captured.

The change of the total cloud field (computed as the linear 30 year trend)  as observed by ISCCP between 1983-2009 is shown in Figure \ref{fig:isccp}.  Although the pattern of cloud cover changes is complex many of the regions of increasing cloud occur over regions with a weak or even decreasing trend of surface temperature (compare with the late period of Figure 2 of SPZ).  This agrees well with the results of Z16 and implies that low-level stability is likely an important factor in the change of global cloud fields.

%To quantify the influence of regional changes in LCC on the TOA radiative budget the long-wave CRE (LWCRE) and short-wave CRE (SWCRE) trends are computed over the two periods of interest.
Global fields for AM4.0 are shown in Figs. 2 and 3 of SPZ and mean values over the windows described in SPZ are given in SM Table 1.  
%The windows used to compute mean values for SM Table 1 were chosen to represent regions of importance highlighted by our results and previous literature.  
%The eastern Pacific window mimics the area of importance discussed in Z16.  We used the definition of the Peruvian and Californian windows from Klein and Hartmann, 1993, expanded them and added them together to measure the LCC and SWCRE trends in the eastern Pacific.  The Indian, south Atlantic, and southern ocean windows were chosen because of the large trends in both the LCC and SWCRE there (see figs. 2 and 3 of SPZ).    
SM Table 1 shows a large increase of LCC trend in the late period in all windows, which coincides to a significant strengthening of the negative SWCRE.  
This should be contrasted with the fact that in the early period all regions have either a decrease in LCC, or only a small increase.  The decreasing LCC of the early period correspond to positive SWCRE.  This confirms the importance of the east Pacific region but shows that the LCC trend in the other regions have an equally large (if not larger) impact on the SWCRE.  
%The changing SWCRE of these 4 regions is consistent with a more sensitive climate during the early period and a less sensitive climate in the  late period.

The linear decomposition of the low-level cloud field given by equation 3 and illustrated with Figure 3.b in SPZ is further examined here.  The results of the decomposition for the latitude ranges $\pm 30^{\circ}$ and $\pm 60^{\circ}$ are tabulated in in SM Table 2.  Shown are the coefficients $\gamma$ and $\beta$ for each model as well as the correlation coefficients between the LCC anomaly from model output compared with the linear approximation of the LCC anomaly using $\gamma$ and $\beta$.  For completeness we also show the time series for the region between $\pm 30^{\circ}$ in Figure 3.

\noindent\textbf{Computational details of the Estimated Inversion Strength}

To compute the EIS following Wood and Bretherton we use 
\begin{equation}
  \rm{EIS}=\rm{LTS}-\Gamma_m^{850}(z_{700}-\rm{LCL})
\end{equation}
with the lifting condensation level (LCL) calculated as $\rm{LCL}=(20+(T_{sfc}-273.15)/5)(1-RH_{sfc})$.

            The potential temperature gradient for moist-adiabatic processes is calculated as
\begin{equation}
    \Gamma_m(T,p)=\frac{g}{c_p} \left[ 1-\frac{1+L_v q_s(T,p)/R_a T}{1+L^2_v q_s(T,p)/c_p R_v T^2} \right].
\end{equation}
            To compute $\Gamma^{850}_m$ we used the temperature on the $850 \rm{hPa}$ pressure level and computed the saturation vapor pressure on the 
            $850 \rm{hPa}$ level using $q_s=610.8 \, \rm{exp}[17.27(T_{850}-273.15)/(T_{850}-35.85)] (\rm{hPa})$.  When the pressure surfaces of interest for these calculations intersect topography 
            the variables we are interested in are not defined.  To account for this, missing vallues of $\rm{T}_{850}$ are set to 50 $K$, $\rm{T}_{sfc}$ is not allowed to be lower than $200$ K, and $\rm{RH}_{sfc}$
            is not allowed to be lower than $5 \%$.

%\clearpage

%Delete all unused file types below. Copy/paste for multiples of each file type as needed.
%\noindent\textbf{Text S1.}
%Type or paste text here. This should be additional explanatory text, such as: extended descriptions of results, full details of models, extended lists of acknowledgements etc.  It should not be additional discussion, analysis, interpretation or critique. It should not be an additional scientific experiment or paper.
%
%Repeat for any additional Supporting Text

%%Enter Data Set, Movie, and Audio captions here
%%EXAMPLE CAPTIONS

%\noindent\textbf{Data Set S1.} %Type or paste caption here.
%upload your dataset(s) to AGU's journal submission site and select "Supporting Information (SI)" as the file type. Following naming %convention: ds01.

%Repeat for any additional Supporting data sets

%\noindent\textbf{Movie S1.} %Type or paste caption here.
%upload your movie(s) to AGU's journal submission site and select, "Supporting Information %(SI)" as the file type. Following naming convention: ms01.

%Repeat any additional Supporting movies

%\noindent\textbf{Audio S1.} %Type or paste caption here.
%upload your audio file(s) to AGU's journal submission site and select "Supporting Information %(SI)" as the file type. Following naming convention: auds01.

%Repeat for any additional Supporting audio files

%%% End of body of article:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Notation section goes here
%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% Or,
%
% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.
%
% Follow these steps:
% 1. Run LaTeX on your LaTeX file.
%
% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
\begin{thebibliography}{6}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
